---
title: "Almond Yield Assignment"
output:
  slidy_presentation:
    highlight: pygments
  html_document: default
  pdf_document: default
  ioslides_presentation:
    highlight: pygments
  beamer_presentation:
    highlight: pygments
---

Submit 
- .R function 
- .RMD
- .html

Your goal is to implement a simple model of almond yield anomaly response to climate
- Inputs: daily times series of minimum, maximum daily temperatures and precipitation
- Outputs: maximum, minimum and mean yield anomoly for a input time series
The Lobell et al. 2006 paper will be the source for the transfer function that you will use in your model; specifically look at the equations in table 2.

Grading 
- a clear diagram that corresponds with your R function (10 pts) 
- inputs and output and parameters shown on the diagram (10 pts)
- correct function implementation (*.R file) (10 pts)
- application of the function to clim.txt (in Rmarkdown file) (10 pts)
- coding practices (clear documentation, informative variables names) (10 pts)

# Draw the Model

# Implement Model in R
- create an R file that has your function definition

```{r}
library(tidyverse)
```


```{r}

yields <- data.frame()

# iterate over the rows to get yields
for(date in 1:nrow(almond_clim)) {
  
  T_min = almond_clim$tmin_c[date]
  P = almond_clim$precip[date]
  
  Y_min = (-0.015*T_min) - (0.0046*(T_min**2)) - (0.07*P) + (0.0043*(P^2)) + 0.28
  
  calculations <- data.frame(Y_min = Y_min)
  
  yields <-  rbind(yields, calculations)
}

max(yields)
min(yields)
mean(yields$Y_min)

```


```{r}
# EQUATION 
almond_func <- function(df){
  
  yields <- data.frame()
  
  # iterate over the rows to get yields
  for(date in 1:nrow(df)) {
    
    T_min = df$tmin_c[date]
    P = df$precip[date]
    
    Y_min = (-0.015*T_min) - (0.0046*(T_min**2)) - (0.07*P) + (0.0043*(P^2)) + 0.28
    
    calculations <- data.frame(Y_min = Y_min)
    
    yields <-  rbind(yields, calculations)
  }
  
  yield <- data.frame(
    Ymax = max(yields),
    Ymin = min(yields),
    Ymean = mean(yields$Y_min)
  )
  
  return(yield)
}

```


# Run the Model
- apply the function to the test climate data (clim.txt)
```{r}

clim <- read.delim('/Users/lunacatalan/Documents/dev/eds230/eds-230-modeling/assignments/clim.txt',
                   header = TRUE,
                   sep = " ")

# select month for almond yields
feb_temp <- clim %>% 
  group_by(year) %>% 
  filter(month == 2) %>% 
  summarize(tmin_c = min(tmin_c),
            tmax_c = max(tmax_c))

# select precipitation for almond yields
jan_precip <- clim %>% 
  group_by(year) %>% 
  filter(month == 1) %>% 
  summarize(precip = sum(precip))

# join the datasets
almond_clim <- left_join(feb_temp, jan_precip)

almond_yield <- almond_func(almond_clim)
```


# Check Work

- the maximum almond yield anomaly should be approximately 1920 ton/acre 
- the lowest almond yield anomaly should be approximately -0.027 ton/acre 
- the mean almond yield anomaly should be approximately 182 ton/acre