---
title: "Predator-Prey Assignment"
authors: "Luna Hershenfeld-Catalan"
date: "2024 May 30th"
---

Consider how you might add hunting  of prey to the predator prey model that we've been using in class

# Part 1: Build this model (e.g add hunting to the lotvmodK.R),

Some requirements/hints for your model

- You should make sure that you don't hunt more prey than exist.

- To ensure that you might also add a minimum prey population (`min_prey`) input that must be met before hunting is allowed.

Note you can make this as simple or as complex as you would like. You could represent hunting in a way that is similar to "harvesting" in the last assignment.  A key challenge is how you might want to define stability? It is up to you but you will need to write a sentence to explain why you chose the measure that you did.  It could be something as simple as maintaining a population above some value 50 years into the future. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
# for sobel
library(sensitivity)
```

```{r include = FALSE}
lotvmodK = function(t, pop, pars) {
  with(as.list(c(pars,pop)), {
    dprey = rprey*(1-prey/K)*prey -  alpha*prey*pred
    dpred = eff*alpha*prey*pred - pmort*pred
    return(list(c(dprey,dpred)))})
}


# initial conditions
currpop=c(prey=1, pred=1)

# set parameter list
pars = c(rprey=0.1, alpha=0.6, eff=0.8,pmort=0.4, K=20)

# times when you want to evaluate
days = seq(from=1,to=500)

# run our differential equation solver
res = ode(func=lotvmodK, y=currpop, times=days, parms=pars)

# rearrange for plotting
resl = as.data.frame(res) %>% pivot_longer(-time, names_to="species", values_to="pop")

# graph both populations over time
p1=ggplot(resl, aes(time, pop, col=species))+geom_line()
p1

# also look at relationships between preditor and prey population and use color for time 
# I will remove the legend here to make it easier to see 
p2 = ggplot(as.data.frame(res), aes(pred, prey, col=(round(time/10))))+geom_point()+theme(legend.position = "none")
p2
p2 = ggplot(as.data.frame(res), aes(pred, prey, col=as.factor(round(time/10))))+geom_point()+theme(legend.position = "none")
p2
ggpubr::ggarrange(p1,p2)
```



```{r}
# initial conditions

lotvmodK_hunt = function(t, pop, pars, min_prey) {
  
  with(as.list(c(pars, pop)), {
    
    #set minimum condition to allow hunting
    if(prey <= min_prey) {
      # subtract the hunted
      rhunt = 0
    }
    
    dprey = rprey*(1-prey/K)*prey -  alpha*prey*pred - rhunt
    dpred = eff*alpha*prey*pred - pmort*pred
    
    return(list(c(dprey,dpred)))
    
  })
}

# initial conditions
currpop=c(prey=1, pred=1)

# set parameter list
pars = c(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, 
         rhunt = 0.001) # create a rate

# times when you want to evaluate
days = seq(from=1,to=500)

# run our differential equation solver
res = ode(y=currpop, 
          times=days, 
          func=lotvmodK_hunt, 
          min_prey = 1,
          parms=pars)

# rearrange for plotting
resl = as.data.frame(res) %>% pivot_longer(-time, names_to="species", values_to="pop")

# graph both populations over time
p1=ggplot(resl, aes(time, 
                    pop, col=species)) + 
  theme_classic() +
  geom_line()

p2 = ggplot(as.data.frame(res), 
            aes(pred, prey, col=(round(time/10)))) + 
  geom_point() + 
  theme_classic() +
  theme(legend.position = "none")

p3 = ggplot(as.data.frame(res), 
            aes(pred, prey, col=as.factor(round(time/10)))) + 
  geom_point() + 
  theme_classic() +
  theme(legend.position = "none")

ggpubr::ggarrange(p1,p3)
```


Part 2

Explore how different hunting levels and different minimum prey populations (before hunting is allowed) are likely to effect the stability of the populations of both predator and prey.

Use this exploration to recommend a hunting target that will be sustainable (e.g leave you with a stable prey and predator population).

It is up to you how you "explore" hunting  - you can simply try different values of the parameters in your hunting model or do it more formally by running your model across a range of values. You could think about parameter interactions

You can assume the following are best guesses of key parameters

rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000

```{r}

np=2000
rhunt= rnorm(mean=0.001, sd=0.001*0.01, n=np)
min_prey = seq(1, np)

pars = list(rhunt=sample(rhunt, 1), 
             min_prey=sample(min_prey, 1), 
             rprey=0.95, alpha=0.01, eff=0.6, pmort=0.4, K=2000)

# run our differential equation solver
res = ode(y=currpop, 
          times=days, 
          func=lotvmodK_hunt, 
          parms=pars)

# rearrange for plotting
resl = as.data.frame(res) %>% pivot_longer(-time, names_to="species", values_to="pop")

# graph both populations over time
p1=ggplot(resl, aes(time, 
                    pop, col=species)) + 
  theme_classic() +
  geom_line()

p3 = ggplot(as.data.frame(res), 
            aes(pred, prey, col=as.factor(round(time/10)))) + 
  geom_point() + 
  theme_classic() +
  theme(legend.position = "none")

ggpubr::ggarrange(p1,p3)
```


```{r}
# generate the sobel parameters
np=2000
rhunt= rnorm(mean=0.001, sd=0.001*0.01, n=np)
min_prey = seq(1, np)
X1 = cbind.data.frame(rhunt=rhunt, min_prey=min_prey)

np=2000
rhunt= rnorm(mean=0.001, sd=0.001*0.01, n=np)
min_prey = seq(1, np)
X2 = cbind.data.frame(rhunt=rhunt, min_prey=min_prey)

# fix any negative values and they are not meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# run sobel model
sens_P = sobolSalt(model = NULL,X1, X2, nboot = 300)

# lets add names 
colnames(sens_P$X) = c("rhunt","min_prey")

simtimes = seq(from=1, to=300)
pars = list(rhunt=sens_P$X[1,"rhunt"], 
             min_prey=sens_P$X[1,"min_prey"], 
             rprey=0.95, alpha=0.01, eff=0.6, pmort=0.4, K=2000)

# initial conditions
currpop=c(prey=1, pred=1)

# run our differential equation solver
res = ode(y=currpop, 
          times=simtimes, 
          func=lotvmodK_hunt, 
         # min_prey = ,
          parms=pars)

res = as.data.frame(res)


```

```{r}
# compute_metrics(result, 100)
p_wrapper = function(rhunt, min_prey, rprey, alpha, eff, pmort, K, currpop, simtimes, lotvmodK_hunt, metricfunc) {
    pars = list(rhunt=rhunt, 
             min_prey=min_prey, 
             rprey=0.95, 
             alpha=0.01, 
             eff=0.6, 
             pmort=0.4, 
             K=2000)
    result = ode(y=currpop, 
          times=simtimes, 
          func=lotvmodK_hunt, 
         # min_prey = ,
          parms=pars)
  # get metrics
  metrics=metricfunc(as.data.frame(result))
  return(metrics)
}

compute_metrics = function(result) {
  
  # select the max result
  maxgrowth = max(result$C)
  
  # set the results that are greater than the threhold
  idx = which(result$prey <= min_prey)[1] # select when hunting is not allowed
  idx = ifelse(is.na(idx), # if no results from greater than threshold
               length(result$C), # then set the index to the length of results 
               idx) # if not keep it
  
  # index the time that the growth passed threshold
  threshyear = result$time[idx]
  
return(list(maxgrowth=maxgrowth, threshyear=threshyear))
}

```



Submit the Rmarkdown that documents

a) your hunting model

b) your exploration (e.g how you tested different hunting levels and how you defined a stability metric

c) provides you estimated sustainable hunting level. and a few sentences justifying this choice 