---
title: "Sobol ODE Assignment"
authors: "Luna Hershenfeld-Catalan"
date: "2024 May 7th"
output: html
---

Task
# 1. Implement model in R

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
# for sobel
library(sensitivity)
```

Create the model as a function:
```{r eval = FALSE}
dforestgrowth = function(Time, C, parms) {
  
  if(C < 50) { # the canopy threshold
    
    dG = parms$r * C
  } else {
    
    dG = parms$r * (1- C/parms$K)
  }
  
  return(list(dG))
}
```


```{r}
# read in the function
source("../../R/dforestgrowth.R")

```


# 2. Run model for 300 years (using ODE solver)

Initial conditions
- initial forest size of 10 kg/C
- canopy closure threshold of 50 kgC
- K = 250 kg C (carrying capacity)
- r= 0.01 (exponential growth rate before before canopy closure)
- g = 2 kg/year (linear growth rate after canopy closure)

Run for 300 years
```{r}
# initial forest size
time = seq(1, 300)
Cinitial = 10
thresh = 50
K = 250
r = 0.01
g = 2

parms = list(K = 250, r = 0.01, g = 2, thresh = 50)
result = ode(y=Cinitial, 
             times=time,
             func=dforestgrowth, # growth function
             parms =parms,
             thresh = thresh)

result <- as.data.frame(result)
```


# 3. Graph Results
```{r}
ggplot(result, aes(time, `1`)) +
  geom_point() +
  theme_classic()
```


# 4. Run sobol global

Sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of C 300 years), varies with these parameters

- pre canopy closure growth rate (r)
- post-canopy closure growth rate (g)
- canopy closure threshold and carrying capacity(K)

Set up the normal distributions for parameters:
```{r message = FALSE, error = FALSE}
# generate the sobel parameters
np=2000
K = rnorm(mean=250, sd=250*0.01, n=np)
r = rnorm(mean=0.01, sd=0.01*0.01, n=np)
g = rnorm(mean = 2, sd = 2*0.01, n = np)
X1 = cbind.data.frame(r=r, K=K, g = g)

# repeat to get our second set of samples
K = rnorm(mean=250, sd=250*0.01, n=np)
r = rnorm(mean=0.01, sd=0.01*0.01, n=np)
g = rnorm(mean = 2, sd = 2*0.01, n = np)
X2 = cbind.data.frame(r=r, K=K, g = g)

# fix any negative values and they are not meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# run sobel model
sens_P = sobolSalt(model = NULL,X1, X2, nboot = 300)

# lets add names 
colnames(sens_P$X) = c("r","K", "g")

```


Get results for 300 years
```{r message = FALSE, error = FALSE}
# gets results for 200 years (evaluating every year)
thresh = 50
simtimes = seq(from=1, to=300)
parms = list(r=sens_P$X[1,"r"], 
             K=sens_P$X[1,"K"], 
             g=sens_P$X[1,"g"])

# run the ODE solver with these conditions
result = ode(y=Cinitial, 
             times=simtimes,
             func=dforestgrowth, # growth function
             parms=parms,
             thresh = thresh)

result = as.data.frame(result)

# rename the column names
colnames(result)=c("time","C")
```


Set up the compute metrics and wrapper functions:
```{r message = FALSE, error = FALSE}

# turn computing our metrics into a function

compute_metrics = function(result, thresh) {
  
  # select the max result
  maxgrowth = max(result$C)
  
  # set the results that are greater than the threhold
  idx = which(result$C > thresh)[1] # select the first year its above threshold
  idx = ifelse(is.na(idx), # if no results from greater than threshold
               length(result$C), # then set the index to the length of results 
               idx) # if not keep it
  
  # index the time that the growth passed threshold
  threshyear = result$time[idx]
  
return(list(maxgrowth=maxgrowth, threshyear=threshyear))
}

# try it on our first parameter set, and look at when it gets to 100
# this works!

# compute_metrics(result, 100)
p_wrapper = function(r, K, g, Cinitial, simtimes, odefunc, metricfunc) {
    parms = list(r=r, K=K, g=g)
    result = ode(y=Cinitial, 
                 times=simtimes, 
                 func=odefunc, 
                 parms=parms,
                 thresh = thresh) 
    colnames(result)=c("time","C")
  # get metrics
  metrics=metricfunc(as.data.frame(result), thresh=100)
  return(metrics)
}

# testing
# this works!
# p_wrapper(r=0.01, K=150, g = 2, Cinitial=3, simtimes=seq(from=1,to=10),
#          odefunc=dforestgrowth, metricfunc=compute_metrics)

```

Apply to everything:
```{r message = FALSE, error = FALSE}
# now use pmap as we did before
all = as.data.frame(sens_P$X) %>% 
  pmap(p_wrapper, 
       Cinitial=Cinitial, 
       simtimes=simtimes,
       odefunc=dforestgrowth,
       metricfunc=compute_metrics)

# extract out results from pmap into a data frame
allres = all %>% 
  map_dfr(`[`,c("maxgrowth","threshyear"))

```



# 5. Graph results of sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T).

```{r message = FALSE, error = FALSE}
# create boxplots
tmp = allres %>% 
  pivot_longer(cols=everything(),
               names_to="metric", 
               values_to="value")

ggplot(tmp, aes(metric, value, col=metric)) + 
  geom_boxplot() + 
  theme_classic()
```


# 6. Discuss meaning of results of simulation (For example think about how what parameters climate change might influence).

The results of these simulations indicate that the max growth year varies mainly between 183 and 186.