---
title: "Sobol Assignment"
authors: "Luna Hershenfeld-Catalan & Sofia Ingersoll"
date: "2024 April 25"
output: html
---
# Part 1

Paragraph on how results of the sensitivity analysis reported on in the paper might contribute to understanding (or prediction) within an environmental problem solving or management context.

# Part 2

Use class model of atmospheric conductance

Cat=vm6.25∗ln(zm−zdz0)2

- Windspeeds v are normally distributed with a `mean of 300 cm/s with a standard deviation of 50 cm/s`

- For `vegetation height` assume that height is somewhere between `3.5 and 5.5 m` (but any value in that range is equally likely)

- For the `kd and k0` parameters you can assume that they are normally distributed with `standard deviation of 1% of their default values`

```{r set_up, message = FALSE}
library(here)
library(tidyverse)

sager <- read.table("../data/sager.txt", header=T) %>% 
  ## add date
  mutate(date = paste(day,month,year, sep="/"))

sager$date <- as.Date(sager$date,"%d/%m/%Y")

#head(sager)

# multiple results - lets say we've run the model for multiple years, each column
# is streamflow for a different parameter set
msage <- read.table("../data/sagerm.txt", header=T)
#head(msage)

# lets say we know the start date from our earlier output
msage$date = sager$date
msage$month = sager$month
msage$year = sager$year
msage$day = sager$day
msage$wy = sager$wy


# how can we plot all results - lets plot water year 1970 otherwise its hard to see
msagel <- msage %>% 
  pivot_longer(cols=!c(date, month, year, day,wy),
               names_to="run",
               values_to="flow")
#head(msage)
```


1. Use the Sobol approach to generate parameter values for the 4 parameters
```{r}

```

2. Run the atmospheric conductance model for these parameters
```{r}

```

3. Plot conductance estimates in a way that accounts for parameter uncertainty
```{r}

```

4. Plot conductance estimates against windspeed use the parameter that is 2nd in terms of total effect on response
```{r}

```

5. Estimate the Sobel Indices for your outut
```{r}

```

6. Comment on what this tells you about how atmospheric conductance and its sensitivity to variation in windspped differs in this setting as compared to the setting that we examined in class where windspeed was lower and less variable and vegetation was taller.



**TAKEN FROM COMPUTE PERFORMANCE METRICS.RMD**

```{r multipel}

# multiple results - lets say we've run the model for multiple years, each column
# is streamflow for a different parameter set
msage = read.table("../Data/sagerm.txt", header=T)

# lets say we know the start date from our earlier output
msage$date = sager$date
head(msage)
msage$month = sager$month
msage$year = sager$year
msage$day = sager$day
msage$wy = sager$wy

# and we still have observed data from above


# how can we plot all results - lets plot water year 1970 otherwise its hard to see
msagel = msage %>% pivot_longer(cols=!c(date, month, year, day,wy), names_to="run", values_to="flow")

p1=ggplot(subset(msagel, wy == 1970), aes(as.Date(date), flow, col=run))+geom_line()+theme(legend.position = "none")
p1
# lets add observed streamflow
p1+geom_line(data=subset(sager, wy == 1970), aes(as.Date(date), obs), size=2, col="black", linetype=2)+labs(y="Streamflow", x="Date")


# compute performance measures for all output
res = msage %>% select(!c("date","month","year","day","wy")) %>%
      map_dbl(nse, o=sager$obs )

head(res)

# another example using our low flow statistics
# use apply to compute for all the data
# using the updated low flow metrics routing that also computed combined metrics

source("../R/compute_lowflowmetrics_all.R")
res = msage %>% select(-date, -month, -day, -year, -wy ) %>% map_df(compute_lowflowmetrics_all, o=sager$obs, month=msage$month, day=msage$day, year=msage$year, wy=msage$wy)


# interesting to look at range of metrics - could use this to decide on
# acceptable values
summary(res)


# graph range of performance measures
resl = res %>% pivot_longer(cols=everything(), names_to="metric", values_to="value")
ggplot(resl, aes(metric, value))+geom_boxplot()+facet_wrap(~metric, scales="free")

# try this
# assign an identifier to each row, use the same identify for columns of original streamflow data
# we can then use that to pick data
res$run = seq(from=1,to=nrow(res))
head(msage)
colnames(msage)=c(res$run, "date","month","year","day","wy")

# best one
best = res[which.max(res$combined),]
msagel  =  msage %>% pivot_longer(cols=!c(date, month, year, day,wy), names_to="run", values_to="flow")
ggplot(subset(msagel, run == best$run), aes(date, flow)) + geom_line()

```

**Taken from lect. 6 on Sobol**
This is for first order indicies
```{r}
# generate two examples of random number from parameter distributions
np=1000
k_o = rnorm(mean=0.1,sd=0.1*0.1, n=np)
k_d = rnorm(mean=0.7, sd=0.7*0.1, n=np)
v = rnorm(mean=250, sd=30, n=np)
height = runif(min=9.5, max=10.5, n=np)
X1 = cbind.data.frame(k_o, k_d, v, height=height)
# repeat sampling
k_o = rnorm(mean=0.1,sd=0.1*0.1, n=np)
k_d = rnorm(mean=0.7, sd=0.7*0.1, n=np)
v = rnorm(mean=250, sd=30, n=np)
height = runif(min=9.5, max=10.5, n=np)
X2 = cbind.data.frame(k_o, k_d, v, height=height)
# there are different versions of sobol functions that have different approaches for estimating
# parameters and indices, we use an approach implemented by jansen
sens_Catm_Sobol = sobolSalt(model = NULL, X1, X2, nboot = 100)
# Take a look at the Sobol generated sensitivity object
# your parameters sets for sensitivity analysis are in X

# run model for all parameter sets
# make sure you give the parameters names
parms = as.data.frame(sens_Catm_Sobol$X)
colnames(parms)= colnames(X1)
res = pmap_dbl(parms, Catm)
sens_Catm_Sobol = sensitivity::tell(sens_Catm_Sobol,res, res.names="ga")
# main effect: partitions variance (main effect without co-variance) - sums approximately to one
sens_Catm_Sobol$S

# useful to add names
row.names(sens_Catm_Sobol$S) = colnames(parms)
sens_Catm_Sobol$S

# total effect - accounts for parameter interactions
row.names(sens_Catm_Sobol$T) = colnames(parms)
sens_Catm_Sobol$T

# Both the main effect and total effect can tell us something about how the parameter influences results
print(sens_Catm_Sobol)

# graph two most sensitive parameters
both = cbind.data.frame(parms, gs=sens_Catm_Sobol$y)
# look at overall gs sensitvity to uncertainty
ggplot(both, aes(x=gs))+geom_histogram()+geom_vline(xintercept=mean(both$gs), col="cyan")

# look at response of conductance to the two most important variables
ggplot(both, aes(k_d,gs, col=k_o))+geom_point()+labs(y="Conductance (mm/s)", x="k_d parameter")

# use second most sensitive parameter (using most important as color)
ggplot(both, aes(k_o,gs, col=k_d))+geom_point()+labs(y="Conductance (mm/s)", x="k_d parameter")
```

*